{\datastorr}: A package for distributing dynamic and versioned data

### 2017-04-24 ms.tex
Add tables listing features, main aims

\title{{\datastorr}: A package for distributing dynamic and versioned
data}
\author{Richard G. FitzJohn$^1$, Daniel S. Falster$^2$, Matthew
  W. Pennell$^3$, and William K. Cornwells$^4$}
\affiliation{
$^1$ Somewhere in the UK\\
$^2$ Everywhere else\\
$^3$ Department of Zoology and Biodiversity Research Centre,
University of British Columbia, Vancouver, B.C. V6T 1Z4 Canada\\

@@

@@

>>>>

\title{{\datastorr}: A package for distributing dynamic and versioned
data}
\author{Richard G. FitzJohn$^1$, Daniel S. Falster$^2$, Matthew
  W. Pennell$^3$, and William K. Cornwell$^2$}
\affiliation{
$^1$ Somewhere in the UK\\
$^2$ Evolution \& Ecology Research Centre, School of Biological, Earth and Environmental Sciences,
University of New South Wales, Sydney, NSW 2052, Australia\\
$^3$ Department of Zoology and Biodiversity Research Centre,
University of British Columbia, Vancouver, B.C. V6T 1Z4 Canada\\

@@

\section{Overview}

Main aims of paper:

\begin{enumerate}
  \item Recognise need for versioned datasets
  \item Provide workflow / infrastructure for people to distribute and access small-medium sized data with minimal setup and cost, by leveraging emerging technologies
\end{enumerate}

% Access point might change, e.g. github goes out of business

% Workflows for dataset assembly: One or two packages? More complex like baad has two packages. Taxonlookup is all in one
% Licensnes? Currently using MIT (taxonlookup), BSD\_2 (baad.data)

@@

\section{Table}

\begin{table}[h!]
\centering
\caption{Alternative frameworks for managing a living database}
{\footnotesize
\vspace{1cm}
  \begin{tabular}{p{5cm}p{5cm}p{5cm}}
  \hline
  \textbf{Feature} & \textbf{Traditional web database} & \textbf{Lightweight versioned data}\\ \hline
   Data storage     & SQL           &  git       \\
   Webserver        & custom        &  Maintained by github.com\\
   Frontend         & Ruby on rails &  datastorr \\
   Ease of setup    & Easy          & Moderate \\ % server, backups, bandwidth
   Data size        & Small-Very large & Up to 1Gb\\
   Cost             & Ongoing       & Free \\
   Bandwidth        & Pro rata      & Managed by github.com\\
   Maintainer skills& Ruby, php     & R, git \\
   User skills      & Web           & R \\
   Versioning       & Hard          & Easy \\
   DOI minting      & Manual        & Automatic \\
  \hline
  \end{tabular}
  }
\label{tab:sql_v_versioneddata}
\end{table}

\newpage

\begin{table}[h!]
\centering
\caption{Main technologies leveraged to store and distribute "Lightweight versioned data".}
{\footnotesize
\vspace{1cm}
  \begin{tabular}{p{5cm}p{10cm}}
  \hline
  \textbf{Technology} & \textbf{Description} \\\hline
   git & Open source version control system used for tracking progressive changes in a set of text files, typically computer code but also data\\
   github.com & A commercial webplatform for sharing, visualising and managing `git' repositories. Includes issue tracking.\\
   R     &  Open source statsitcial and data processing language. \\
   CRAN  &  Open source repository of packages for the R language \\
   datastorr & R package used to fetch github releases. \\
   \hline
  \end{tabular}
  }
\label{tab:technologies}
\end{table}

\newpage

\begin{table}[h!]
\centering
\caption{Groups of users and their requirements}
{\footnotesize
\vspace{1cm}
  \begin{tabular}{p{2cm}p{5cm}p{7cm}}
  \hline
  \textbf{Group} & \textbf{Goal} & \textbf{Requirements} \\ \hline
  Maintainer & Create and distribute versioned datasets & Low technical overhead \\
    & & Easy workflow for releasing new versions \\
    & & Long term preservation \\
    & & Easy to crowd-source contributions \\
    & & Low cost \\
    & & Low maintenance \\
  Contributor & Contribute to future versions & Add new data \\
    & & Report errors in existing data  \\
  Users (all) & Easy access to releases & Introduction \& overview \\
    & & Long term stability \\
    & & Clear path for user to become contributor \\
  Users (computational, machines) & Scripted access to releases & Easy installation\\
    & & Long term stability \\
  \hline
  \end{tabular}
  }
\label{tab:user_requirements}
\end{table}

\newpage

\begin{table}[h!]
\centering
\caption{Alternative platforms for minting DOIs from github repositories}
{\footnotesize
\vspace{1cm}
  \begin{tabular}{p{3cm}p{8cm}}
  \hline
  \textbf{Provider} & \textbf{Features} \\ \hline
  \multicolumn{2}{l}{Zenodo: zenodo.org}\\
    & Automated DOI for each github release \\
    & Funded and maintained by CERN and OpenAIRE for long term preservation \\
    & Only includes files embedded in git repo, not built  by code\\
  \multicolumn{2}{l}{Figshare: figshare.com}\\
    & Builds data package based on github release \\
    & Allows you to upload additional files, e.g. datasets built by code \\
    & Integrated with several journals, e.g. Nature family, PNAS \\
    & Some limited potential for versioned DOIs\\
  \hline
  \end{tabular}
  }
\label{tab:doi_minting}
\end{table}

\newpage
